Membrane-localized Keratin-14 Promotes Invasion\thanks{Supported by organization x.}

### 2019-01-12 docs/01_paper/back-up.tex
Added more figures. Finished methods.

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Membrane-localized Keratin-14 Promotes Invasion} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Yohannes Tsehay\textsuperscript{1},
Hildur Knutsodottir\textsuperscript{1},
Andrei Kucharavy\textsuperscript{1\textcurrency},
Veena Padmanaban\textsuperscript{2},
Joel Bader\textsuperscript{1*}
%with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} High-Throughout Biology Center and Department of Biomedical Engineering, Johns Hopkins University, Baltimore, MD, USA
\\
\textbf{2} Center for Cell Dynamics and Department of Cell Biology, Johns Hopkins University School of Medicine, Baltimore, MD, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Metastasis is the main predictor of outcome for patients with cancer, yet the molecular mechanism driving invasive phenotypes is not well under-stood, hindering development of assays and therapeutics for this stage of cancer progression. In this study, we characterize a molecular biomarker for invasion in greater detail, using an automated computer model. The model quantifies the invasive potential of tumors, allowing for quantification of even minor contributions to an invasive phenotype from individual genes, and can potentially be used to establish interactions among genes. Keratin 14 (K14) is a known molecular biomarker that is correlated with poor outcomes in breast cancer patients. We apply our automated computer model, which was built in Python, to test the hypothesis that K14 is directly correlated with automatically generated invasion scores, independent of tumor size. We used two different transgenic mice as our breast cancer tumor models. Organoids were generated from these mice, and each organoid had a corresponding differential interference contrast image, K14 image, and a boundary around the organoid drawn manually which was used to generate an invasion score for the organoid using our automated system. The parameters assessed include the following: entire K14 expression, peripheral K14 expression (edges of the organoid), central K14 expression (center of the organoid), and organoid size. Peripheral K14 expression showed the strongest correlation with invasion scores for both types of transgenic mice, suggesting that K14 expression in cells located in the periphery may be an important marker of invasion. This relationship was more prominent in one mouse model than the other, indicating that additional factors play a role in invasion and dissemination. In future studies, we plan to perform a similar analysis using human organoids generated from breast tumors.

@@

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
%Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

Alternative title: Morphological features of organoid invasion are correlated with membrane-localized Keratin-14 expression

@@

 A total of 4 PyMT and 4 C31T mice were used to generate a total of \{\} organoids, which were then cultured in collagen I gels for 5 days. At day 5, they were imaged, an average of 16 per sample. Differential interference contrast (DIC) microscopy yielded 1040$\times$1388 pixel images. Microscope optics defined a resolution of 0.51190476 $\mu$m per pixel, corresponding to a 532.4$\times$710.5 $\mu$m field of view. Images were manually tracked in {\sc ImageJ} \cite{Schneider:2012ui} to define the organoid boundaries as pairs of points $\{x_v,y_v\}$ for $v \in [0,V-1]$. The total number of points $V$ depended on the manual tracking, and the spacing between adjacent points was similarly variable.

@@

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Morphological Features of Organoid Invasion are Correlated with Membrane-Localized Keratin-14 Expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Yohannes Tsehay\textsuperscript{1},
Hildur Knutsodottir\textsuperscript{1},
Andrei Kucharavy\textsuperscript{1\textcurrency},
Veena Padmanaban\textsuperscript{2},
Joel Bader\textsuperscript{1*}
%with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} High-Throughout Biology Center and Department of Biomedical Engineering, Johns Hopkins University, Baltimore, MD, USA
\\
\textbf{2} Center for Cell Dynamics and Department of Cell Biology, Johns Hopkins University School of Medicine, Baltimore, MD, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Metastasis is the main predictor of outcome for patients with cancer, yet the molecular mechanism driving invasive phenotypes is not well understood, hindering development of assays and therapeutics for this stage of cancer progression. In this study, we characterize a molecular biomarker for invasion in greater detail, using an automated computer model. The model quantifies the invasive potential of tumors, allowing for quantification of even minor contributions to an invasive phenotype from individual genes, and can potentially be used to establish interactions among genes. Keratin 14 (K14) is a known molecular biomarker that is correlated with poor outcomes in breast cancer patients. We apply our automated computer model, which was built in Python, to test the hypothesis that K14 is directly correlated with automatically generated invasion scores, independent of tumor size. We used two different transgenic mice as our breast cancer tumor models. Organoids were generated from these mice, and each organoid had a corresponding differential interference contrast image, K14 image, and a boundary around the organoid drawn manually which was used to generate an invasion score for the organoid using our automated system. The parameters assessed include the following: entire K14 expression, peripheral K14 expression (edges of the organoid), central K14 expression (center of the organoid), and organoid size. Peripheral K14 expression showed the strongest correlation with invasion scores for both types of transgenic mice, suggesting that K14 expression in cells located in the periphery may be an important marker of invasion. This relationship was more prominent in one mouse model than the other, indicating that additional factors play a role in invasion and dissemination. In future studies, we plan to perform a similar analysis using human organoids generated from breast tumors.

@@

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Metastasis is the main predictor of outcome for patients with cancer. The relative 5 year survival rate in patients with breast cancer in situ, regional spread, and distant metastasis are $99\%$, $85\%$, and $26\%$ respectively\cite{Siegel:2017cs}. The survival rates for all stages of breast cancer have improved substantially, but much of the improvement is seen among patients with regional stage cancer with a 5-year relative survival rate increase from $68\%$ to $89\%$ in white women and $55\%$ to $81\%$ in black women. The improvements seen are largely due to adjuvent antineoplasitc, hormonal, and targeted therapy, and early tumor detection with mammograms \cite{DeSantis:2017BC}. Secular trends of tumor size within each stage of the disease (local, regional, and distant) has been decreasing since 1975 due to early detection of tumors\cite{Elkin:2005cts}, yet tumor size or even other tumor characteristics such as estrogen receptor status do not explain this positive trend in survival rates especially in patients younger than 70 years of age \cite{Park:2015ibc}. Thus, it is important to further our understanding of the molecular mechanism of tumor invasion and dissemination to support the development of therapeutics that can improve outcomes for patients with invasive carcinoma and establish biomarkers that accurately predict patient prognosis.

Three-dimentional organotypic culture techniques can be used to model disease processes in vitro. Tissue samples can be harvested from human or murine donors and processed by mechanical disruption, enzymatic digestion, and centrifugation into tissue fragments known as organoids which are explanted into a commercial extracellular matrix such as Matrigel or collagen I\cite {Shamir:2014td}. This experimental technique combined with transgenic mice tissue donors allows for a series of experiments that can elucidate the relationship between gene expression and phenotypic behavior of mammalian tissue. For example, MMTV-PyMT and C3(1)-TAg are two different transgenic mice that develop metastatic carcinoma of the breast\cite{Maroulakou:1994pm,Guy:1992im}. These mice have provided a supply of organoids used to study the invasive process of breast tumors\cite{Shamir:2014td,Cheung:2013ci,Cheung:2016pcb}.

The various clinical and organotypic cell culturing studies have identified Keratin 14 (K14) as an important biomarker. K14 is a structural protein normally expressed in basal epithelial cells or myoepithelial cells. Its expression in luminal glandular cells, which seldom express K14, has been linked to decreased relapse-free survival and overall poor outcomes for patients with breast cancer\cite{Gusterson:2005bc,Laakso:2006blc}. Organotypic cell culture experiments have demonstrated that breast tumors are heterogeneous, and it is specifically the K14 expressing cells that initiate the invasion process. Furthermore, locally disseminated cells, circulating tumor cells and those that have seeded distant organs all frequently express K14\cite{Cheung:2013ci,Cheung:2016pcb}. However, how K14 promotes invasion, whether it be through cell signaling pathways, alterations in the cytoskeleton, or changes in a cell's relationship with other neighboring cells, remains unclear.

Spectral Power

Objective and hypothesis

%Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

%Alternative title: Morphological features of organoid invasion are correlated with membrane-localized Keratin-14 expression

@@

 A total of 4 PyMT and 4 C31T mice were used to generate a total of $1577$ ($880$ PyMT and $697$ C31T) organoids, which were then cultured in collagen I gels for 5 days. At day 5, they were imaged. Differential interference contrast (DIC) microscopy yielded 1040$\times$1388 pixel images. Microscope optics defined a resolution of 0.51190476 $\mu$m per pixel, corresponding to a 532.4$\times$710.5 $\mu$m field of view. Images were manually tracked in {\sc ImageJ} \cite{Schneider:2012ui} to define the organoid boundaries as pairs of points $\{x_v,y_v\}$ for $v \in [0,V-1]$. The total number of points $V$ depended on the manual tracking, and the spacing between adjacent points was similarly variable.

@@

\bibitem{Siegel:2017cs}
Siegel, Rebecca L., Kimberly D. Miller, and Ahmedin Jemal. 
\newblock {"Cancer statistics, 2017."}
\newblock {CA: a cancer journal for clinicians 67.1 (2017): 7-30.}

\bibitem{DeSantis:2017BC}
DeSantis, Carol E., et al.
\newblock {"Breast cancer statistics, 2017, racial disparity in mortality by state."}
\newblock{CA: a cancer journal for clinicians 67.6 (2017): 439-448.}

\bibitem{Elkin:2005cts}
Elkin, Elena B., et al. 
\newblock {"The effect of changes in tumor size on breast carcinoma survival in the US: 1975–1999."}
\newblock {Cancer: Interdisciplinary International Journal of the American Cancer Society 104.6 (2005): 1149-1157.}

\bibitem{Park:2015ibc}
Park, Ju-Hyun, William F. Anderson, and Mitchell H. Gail. 
\newblock {"Improvements in US breast cancer survival and proportion explained by tumor size and estrogen-receptor status."}
\newblock {Journal of Clinical Oncology 33.26 (2015): 2870.}

\bibitem {Shamir:2014td}
Shamir, Eliah R., and Andrew J. Ewald.
\newblock {"Three-dimensional organotypic culture: experimental models of mammalian biology and disease."}
\newblock {Nature reviews Molecular cell biology 15.10 (2014): 647.}

\bibitem {Maroulakou:1994pm}
Maroulakou, Ioanna G., et al. 
\newblock{"Prostate and mammary adenocarcinoma in transgenic mice carrying a rat C3 (1) simian virus 40 large tumor antigen fusion gene."} \newblock {Proceedings of the National Academy of Sciences 91.23 (1994): 11236-11240.}

\bibitem {Guy:1992im}
Guy, Chantale T., R. D. Cardiff, and William J. Muller. 
\newblock {"Induction of mammary tumors by expression of polyomavirus middle T oncogene: a transgenic mouse model for metastatic disease."} \newblock {Molecular and cellular biology 12.3 (1992): 954-961.}

\bibitem {Cheung:2013ci}
Cheung, Kevin J., et al. 
\newblock {"Collective invasion in breast cancer requires a conserved basal epithelial program."}
\newblock {Cell 155.7 (2013): 1639-1651.}

\bibitem {Cheung:2016pcb}
Cheung, Kevin J., et al.
\newblock{"Polyclonal breast cancer metastases arise from collective dissemination of keratin 14-expressing tumor cell clusters."}
\newblock {Proceedings of the National Academy of Sciences 113.7 (2016): E854-E863.}

\bibitem{Gusterson:2005bc}
Gusterson, Barry A., et al. 
\newblock {"Basal cytokeratins and their relationship to the cellular origin and functional classification of breast cancer."}
\newblock {Breast Cancer Research 7.4 (2005): 143.}

\bibitem{Laakso:2006blc}
Laakso, Mervi, et al. 
\newblock {"Basoluminal carcinoma: a new biologically and prognostically distinct entity between basal and luminal breast cancer."} \newblock {Clinical Cancer Research 12.14 (2006): 4185-4191.}
